% this will be the acknowledgement section. 

... Mark Stacey who gave me the freedom to develop my research through trial and error in a beautiful location,.. 
\emph{and trusted me even when that error involved burying instruments again and again...}. Mark 

% First: Mark, thesis/exam committee: Tina Chow, John Largier, Stephanie Carlson, Jim Hunt.

... and Tina Chow, John Largier, Stephanie Carlson, and Jim Hunt, who have... 
my committee who kept me on my toes by asking the opposite of what I felt I had prepared for in oral exams, and were excellent instructors, mentors, etc... 

My committee, Tina Chow, John Largier, Stephanie Carlson, and Jim Hunt, served to... 

% Second: Officemates, colleageus (Bowen, Audric, Susan, Wayne, Rusty, Rudi, Ian, Lauren, Kevin, Maddie, Rachel, Margaret, Cristina, Maureen)... Daniella and Case here. 
I was lucky to be surrounded by brilliant colleagues and wonderful friends in the environmental fluid mechanics group.Among them, Susan Willis, Rusty Holleman, Wayne Wagner, Ian Tse, Audric Collignon, Lauren Goodfriend, Bowen Zhao, Maureen Downing-Kunz, Margaret Byron, Cristina Poindexter, Dave Wiersema, Maddie Foster-Martinez, and Rachel Allen were there help figure out what it means to be a graduate student.  Whether through long hours of field work, over coffee or beers, or grappling with ... the community in O'Brien Hall... \emph{is characteristic of what scientfic }


... and colleagues through graduate school Daniella Rempe and Case ven Genuchten. 

% Third: Field help
Measurements in the Pescadero estuary would not have been possible with out field assistance from a long list of people. Along with my promises of a day at the beach, Mother Nature often delivered cold water, buried moorings, wind, and rain. I am especially thankful for the willingness of several people to spend hours digging in the sand below cold moving water... Not included in this dissertation is our finding that it is impossible to dig a milk crate sized hole in the sand below 40 cm of water without it filling in... but concerted efforts were made nonetheless.  For this help and more, I thank: Cristina Poindexter, Audric Collignon, Andreas Brand (without whom I may still be trying to figure out how to build and deploy mooorings), Susan Willis, Rusty Holleman, Wayne Wagner, Rocky Geyer, Jordon Hemingway, Kevin Hsu, Rudi Schuech, Ian Tse, Mark Stacey, Felix Sturm, Bowen Zhao, Lauren Goodfriend, Andy Torkelson, Jason Simon, Dave Weirsema, DJ Gaker, Eric Huber, Dane Behrens, Nate Butler, Justin Jasper, Gabrielle Boisrame, Arthur Wiedmer, Gabriele Bellani, Hugh Wagner, Julia Schoen, Ryan Sanders, and those friends and collegues whose assistance I failed to write down in my field notebook. 

This work was permited by California State Parks.

% Fourth: coop ...   2009-2013 residents of the 
While living in Berkeley... 
I would like to acknowledge the Hillegass-Parker Cooperative house for introducing me to great minds and great friends, especially Matt Tucker-Simmons, Katie Adamides, and Tim Ruckle.... and to the members and teachers in the Berkeley Art Studio, especially Hue Yang and Jim Newton...

... the art studio gave me a much needed sense of accomplishment that was hard to find in the middle of my PhD work. 

% Funding
This research and dissertation would not have been possible without the support of a National Science Foundation Graduate Research Fellowship (DGE 1106400).

% Fifth: 
Finally, I want to acknowledge the unlikely coupling of the California prison system and its environmental management.