\documentclass{ucbthesis}
\usepackage{graphicx} % ADDED BY AUDRIC COLLIGNON 06/07/2012: insert figures
\usepackage{rotating} % ADDED BY AUDRIC COLLIGNON 07/23/2012: tables in landscape mode
\usepackage{lscape} % ADDED BY AUDRIC COLLIGNON 07/23/2012: table in landscape mode
\RequirePackage{lineno} % ADDED BY AUDRIC COLLIGNON 07/23/2012: LINE NUMBERING PACKAGE - REMOVE FROM FINAL VERSION
\usepackage{textcomp}  %Added 07/11/2014 by M. Williams.. to get copyright symbol in ch2
\usepackage{subcaption}
\usepackage{epstopdf}
\epstopdfDeclareGraphicsRule{.tif}{png}{.png}{convert #1 \OutputFile}  % for dealing with tif
%\AppendGraphicsExtensions{.tif}
\usepackage{epsfig}

% color... 
% \usepackage{xcolor}
% \color{blue}


\usepackage[style=authoryear, maxcitenames=2, maxbibnames=99, url=false, clearlang=true, isbn=false, firstinits]{biblatex} %STYLE OPTION ADDED BY AUDRIC COLLIGNON 07/26/2012: use author (year) citation style
\renewbibmacro{in:}{} %ADDED BY AUDRIC COLLIGNON 07/27/2012: remove In: in front of journal name in bibliography
\urlstyle{same}


% \renewbibmacro*{volume+number+eid}{%
%   \printfield{volume}%
% %  \setunit*{\adddot}% DELETED
%   \setunit*{\addnbspace}% NEW (optional); there's also \addnbthinspace
%   \printfield{number}%
%   \setunit{\addcomma\space}%
%   \printfield{eid}}
% \DeclareFieldFormat[article]{number}{\mkbibparens{#1}}

% This gets volume.number to be volume(number)
\usepackage{xpatch}
% No dot before number of articles
\xpatchbibmacro{volume+number+eid}{%
  \setunit*{\adddot}%
}{%
}{}{}
% Number of articles in parentheses
\DeclareFieldFormat[article]{number}{\mkbibparens{#1}}




\setcounter{tocdepth}{3} % 3 will put up to subsubsections into the TOC
\setcounter{secnumdepth}{2} % 2 labels sections, 3 labels subsections. 



% If the Grad. Division insists that the first paragraph of a section
% be indented (like the others), then include this line:
% \usepackage{indentfirst}
%\bibliographystyle{plain}
\bibliography{EntireZoteroLibrary}
\bibliography{research}
\bibliography{thesis_refs_added}

\AtEveryBibitem{\clearlist{language}} % clears language
\AtEveryBibitem{\clearfield{note}}    % clears notes
\AtEveryBibitem{\clearfield{month}}
\AtEveryCitekey{\clearfield{month}}



\begin{document}

% \linenumbers % ADDED BY AUDRIC COLLIGNON 07/23/2012: LINE NUMBERING COMMAND - REMOVE FROM FINAL VERSION

% Declarations for Front Matter

\title{Hydrodynamics and Salt Dispersion in Intermittently Closed Bar-Built Estuaries}
\author{Megan E. Williams}
\degreesemester{Fall}
\degreeyear{2014}
\degree{Doctor of Philosophy}
\chair{Professor Mark T. Stacey}
\othermembers{Associate Professor Fotini Katapodes Chow\\Professor John L. Largier\\Assistant Professor Stephanie M. Carlson}
\numberofmembers{4}
\field{Engineering - Civil and Environmental Engineering}
\campus{Berkeley}

\maketitle    % INCLUDE IN FINAL VERSION
% \approvalpage %!!! REMOVE IN FINAL VERSION !!!
\copyrightpage  % INCLUDE IN FINAL VERSION

\include{abstract/abstract}  % PUT THIS BACK WHEN YOU'VE WRITTEN AN ABSTRACT.

\begin{frontmatter}

% \begin{dedication}
% \null\vfil
% \begin{center}
% To...
% \end{center} 
% \vfil\null
% \end{dedication}

% % PUT ALL OF THESE BACK LATER: ---------------------
\tableofcontents
\clearpage
\listoffigures
% \clearpage
\listoftables
% \clearpage

\begin{acknowledgements}
% write acknowledgements here... include committee, officemates, massive list of field help, ... 
% this will be the acknowledgement section. 

... Mark Stacey who gave me the freedom to develop my research through trial and error in a beautiful location, and showed me that .. 

... and passed on his genuine excitement for our field... 

\emph{and trusted me even when that error involved burying instruments again and again...}. Mark 


My committee, Tina Chow, John Largier, Stephanie Carlson, and Jim Hunt, were excellent teachers and mentors, and kept me on my toes by asking oral exam questions that were the opposite of those for which I prepared.

I was lucky to be surrounded by brilliant colleagues and wonderful friends in the environmental fluid mechanics group. Among them, Susan Willis, Rusty Holleman, Wayne Wagner, Ian Tse, Audric Collignon, Lauren Goodfriend, Bowen Zhou, Maureen Downing-Kunz, Margaret Byron, Cristina Poindexter, Dave Wiersema, Maddie Foster-Martinez, and Rachel Allen were there help figure out what it means to be a graduate student.  Whether through long hours of field work, grappling with the physical meaning behind a strange piece of data, or the consumption of ample coffee and beer, the community in O'Brien Hall created an invaluable environment in which to pursue doctoral research over five years. 

....Furthermore, Daniella Rempe and Case van Genuchten 
... and colleagues through graduate school Daniella Rempe and Case ven Genuchten. 

% Third: Field help
Measurements in the Pescadero estuary would not have been possible with out field assistance from a long list of people. Along with my promises of a day at the beach, Mother Nature often delivered cold water, wind, and rain. I am especially thankful for those people who spent hours digging in the sand below cold water to free buried moorings. For this help and more, I thank Cristina Poindexter, Audric Collignon, Andreas Brand (without whom I may still be trying to figure out how to build and deploy mooorings), Susan Willis, Rusty Holleman, Wayne Wagner, Rocky Geyer, Jordon Hemingway, Kevin Hsu, Rudi Schuech, Ian Tse, Mark Stacey, Felix Sturm, Bowen Zhou, Lauren Goodfriend, Andy Torkelson, Jason Simon, Dave Weirsema, DJ Gaker, Eric Huber, Dane Behrens, Nate Butler, Justin Jasper, Gabrielle Boisrame, Arthur Wiedmer, Gabriele Bellani, Hugh Wagner, Julia Schoen, Ryan Sanders, and those friends and collegues whose assistance I failed to write down in my field notebook. Furthermore, California State Parks provided permits for fieldwork in the Pescadero estuary.  

% Fourth: coop ...   2009-2013 residents of the 
The Hillegass-Parker Cooperative house introduced me to great minds and great friends, and I thank all of the residents there from 2009 - 2013, especially Matt Tucker-Simmons, Katie Adamides, and Tim Ruckle, for being my family and making the house into my home. The Berkeley Art Studio, and members and instructors in the ceramics studio there also contributed to my graduate school education. 

% Funding
This research and dissertation would not have been possible without the support of a National Science Foundation Graduate Research Fellowship (DGE 1106400).

% Fifth: 
Finally, I want to acknowledge the unlikely coupling of the California prison system and its environmental management. In multiple occasions

\end{acknowledgements}

\end{frontmatter}
% % PUT ALL OF THESE BACK LATER: ---------------------


\pagestyle{headings}  % !!! I think this has to stay out at the end?!!!

%\include{chapter0/chapter0}

\include{chapter1/chapter1}

\include{chapter2/chapter2}

\include{chapter3/chapter3}

\include{chapter4/chapter4}

\include{chapter5/chapter5}

\include{chapter6/chapter6}

% \appendix


\printbibliography


\end{document}
