\newcommand{\undefinedpagestyle}{} % added by Audric Collignon 06/08/2012 to get rid of an error message when compiling

\begin{abstract}

Shallow, bar-built estuaries on wave-dominated coasts in Mediterranean climates experience an intermittent connection to the ocean. Their inlets may completely close as a result of nearshore sand transport, but even the open condition these inlets remain constricted. Extensive field measurements in the highly salt-stratified Pescadero estuary in Northern California were made to elucidate hydrodynamic processes in these ecologically important yet understudied systems. 
Here we present unique detailed observations and a framework for understanding dynamics in small, shallow bar-built estuaries. 

Observations show that the shallow mouth causes these estuaries to experience discontinuous tidal forcing. While the ocean and estuary are fully connected with near equal water levels, tidal velocities are slow but infragravity motions in the nearshore cause large velocity oscillations. As the ocean tide falls, infragravity forcing is cut off because the estuarine mouth is perched above the low tide ocean water level, and ebbing velocities are set by bed friction. Observations reveal this oscillation between ocean-forced conditions and frictionally-controlled conditions characterizes and sets the hydrodynamics and salt dynamics in these estuaries. Additional wave setup of the lagoon water level emphasizes the dependence of these estuaries on nearshore ocean conditions, but the diurnal or semidiurnal retreat of the ocean below the mouth cuts off this nearshore influence. The salt field responds to this discontinuous forcing, being transported upstream on the flood and becoming trapped in deep pools of the estuary on ebb.

Longitudinal salt dispersion within the Pescadero estuary is calculated using velocity and salinity records. Dispersion of high salinity water was found to be an order of magnitude lower than dispersion of low salinity water. Vertical salinity measurements indicate that on the large ebb tide, high surface salinities rapidly relax, and a second pulse of salty water follows the relaxation.  This pulse may induce vertical mixing in the fresher upper water column, and incites the high dispersion seen among low salinities. Upstream processes are implicated in trapping lower salinity water, while higher salinity water is bathymetrically restricted from marsh or shallow channel trapping...  . 




Finally, observations were being made in the Pescadero estuary in March 2011 when the tsunami generated by the Tohok$\overline{\mathrm{u}}$ earthquake arrived to the California coast. Measurements show that the presence of a tsunami-period signature in this shallow, bar-built estuary was modulated by tidal stage and that the tsunami-induced rise and fall of the ocean water level modulated the velocities seen within the estuary. The initial arrival of waves due to the tsunami caused some flushing of the estuary. Later, high velocities induced more mixing than typically seen in this estuary. After several days sand transport nearly choked closed the mouth of the estuary. These observations point to mechanisms which were likely important during the 2011 Tohoku tsunami in many Eastern Pacific estuaries. 

\end{abstract}