\chapter{Introduction}
\label{chIntro}

Along the world's coastlines, rivers and streams drain all the planet's land area through estuaries.  These estuaries are in many cases nursery environments XXX, but also with large population centers around often natural harbor XXX, and with international shipping, and ETC, estuaries are also polluted and invaded \emph{infested, overrun, occupied} by \emph{invasive species}. 

Physics of estuaries is \emph{interesting} because where salty ocean water and fresh river water meet, density differences XXXXX. Add to that XXXXX forcing by tidal flow, changing riverine discharge, wind forcing, and in large estuaries the effect of Earth's rotation, and the understanding of estuarine physics is a complicated affair. Characterizing conditions for \emph{bion00b thing}, understanding contaminant transport, ETC, motivates the XXXX study of these environments. 

Within the state of California on North America's west coast, the San Francisco Bay drains the Sacramento-San Joaquin watershed, consisting of the California Central Valley and western slope of the Sierra Nevada. California's Coastal Range lies east of the Central Valley, and these watersheds drain the western side of California through small bar-built estuaries.

Like larger estuaries, these small systems are historically \emph{(steller places)} for fishes, host migratory water birds, and generally support ecosystem function \emph{(that sounds like an ecology word)}. Coho salmon and steelhead trout pass through California's small estuaries en route to spawn in the headwaters, and those fish who spend time in \emph{(something)} rich lagoons perched behind sand barriers grow faster and are more likely to survive their \emph{(part of?)} life in the Pacific Ocean \parencite{hayes_steelhead_2008, bond_marine_2008}. 

Bar-built estuaries are formed when \emph{(XXXXX)}. In California, these estuaries are small and have an intermittent connection to the ocean. The sandy inlet of these estuaries may close for days, seasonally, or may only temporarily open with large flood events. 

%\empH{they exist worldwide:}
Intermittently closed estuaries exist worldwide, primarily on wave-dominated coasts with seasonal rainfall. Those in South Africa, where have been named temporary open/closed estuaries (TOCE), and Australia, where they are called intermittently closed/open lakes and lagoons (ICOLL), have received the most scientific attention. Estuaries with temporarily closed inlets also exist in Chile \parencite{dussaillant_water_2009}, Brazil \parencite{suzuki_effects_1998}, Spain \parencite{moreno_morphodynamics_2010}, Portugal \parencite{fortunato_morphological_2014}, New Zealand \parencite{schallenberg_contrasting_2010}, India and Sri Lanka \parencite{ranasinghe_seasonal_2003}, and likely elsewhere.Estuarine processes important in intermittently closed estuaries may also be relevant to small bar-built estuaries that do not experience inlet closure. 

Traditional estuarine hydrodynamics research has focused on much larger estuaries, so despite the prevalence of intermittently closed estuaries, they remain understudied.  Several factors may play into this. These small estuaries are not often navigable by large ships, limiting their economic importance. Estuarine physics research is usually within the realm of physical oceanography or environmental engineering, and the small size of intermittently closed estuaries contrasts with the scales typically studied by physical oceanographers.

Short of falsely saying that there is a complete dearth of geophysical research on intermittently closed estuaries, some important work is highlighted below. Many studies focus on inlet morphodynamics, especially on closure mechanisms, though the ability to predict mouth closure is still elusive. Some work has focused on hydrodynamics. A brief description of our current understanding follows. When possible, examples from California are cited.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Inlet morphodynamics}

A full review of sand transport in inlets would overwhelm the intended purpose of this introduction, but I will attempt a quick summary relevant to this dissertation research. Research on morphodynamics tends to focus on inlet closing and conditions setting this closure. A small set of papers addresses how the sand berm grows after inlet closing, and at least a few experimental papers have addressed how the sand berm is eroded upon breach.

\paragraph{closure papers} 
Ranasinghe and Pattiaratchi \parencite*{ranasinghe_seasonal_2003} define two mechanisms for the sand transport required for inlet closure: (1) by longshore currents, or (2) by sandbars migrating on shore (cross-shore currents). Either of these mechanisms interacts  with tidal and riverine flow at the mouth. Tidal flow may scour or XXXXX the inlet as XXXXX\emph{asymmetry}, although observations suggest this is negligible compared to the \emph{things i just said}. Riverine flow will scour the inlet.  Davidson et al. \parencite*{davidson_simple_2009} add to these two processes the downslope diffusion of sediment. The mechanism responsible for mouth closure may be seasonally varying for the same estuary. Examining a X year dataset, Behrens et al. \parencite*{behrens_episodic_2013} found closure of the Russian River mouth likely to be due to cross-shore transport during the summer and early fall, but due to longshore transport during other seasons.

Some examples of conditions for mouth closure follows. 

In a tidal marsh in San Francisco's Crissy Field, located inside of the Golden Gate on the San Francisco Bay, the sandy inlet closed either after a period of large waves (typically from the northwest), but also was seen to close during a small wave climate with long period swell from the south or southwest \parencite{hanes_waves_2011}. 

Closure of the Russian River estuary varies on seasonal timescales. During low flow conditions characteristic of summer Mediterranean climates, the probability of inlet closure increased with ocean wave height \parencite{behrens_episodic_2013}. \emph{River conditions set \# of days closed. mention extensive dataset}

Closure papers: \parencite{ranasinghe_seasonal_2003, elwany_opening_1998}

\paragraph{berm increase}
After the estuarine inlet has closed, swash overwash is responsible for berm development \parencite{baldock_morphodynamic_2008}. Two modes have been proposed to grow the berm: vertical berm growth due to X and horizontal berm growth due to X \parencite{weir_beach_2006}. 

Laudier et al. \parencite*{laudier_measured_2011} measured the lagoon water level at the Carmel River lagoon in Central California and ... \emph{only talked about wave overwash... i think no connection to berm development}


In many cases, closed inlet opening is artificial (e.g. \cite{fortunato_morphological_2014, behrens_episodic_2013}). Stretch and Parkinson investigated beach dynamics by building an experimental lagoon, artifically cutting an inlet, and observing...
Opening/breach papers: \parencite{rich_hydrologic_2013, parkinson_breaching_2007, stretch_breaching_2006, elwany_opening_1998, fortunato_morphological_2014}
Stretch and Parkinson built an experimental lagoon, cut a small inlet, and made measurements of the resulting breach (\cite*{stretch_breaching_2006}; \cite{parkinson_breaching_2007}).



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Intermittently open estuarine hydrodynamics}

Historically, hydrodynamic observations made in South African then Australian intermittently open estuaries set the framework for understanding general salt dynamics and circulation within these estuaries. Work in South African bar-built estuaries defined the general circulation: salt water enters the estuary as a tidal intrusion front owing to the constricted/sill-like inlet, flows upstreawm as a density current, and salt removal is by shear entrainment into fresher (outflowing) surface waters \parencite{largier_stratified_1992}. Field measurements in Wilson Inlet in southwestern Australia show a similar salinity intrusion and trapped salt water in the open state \parencite{ranasinghe_circulation_1999}. Wilson Inlet is a large coastal lagoon, so \emph(coriolis, wind thing), but some similarities to \emph{(what?)} suggest some processes remain constant across length scales. 

In the closed-mouth state of intermittently open estuaries, stratification and circulation are most often studied in the context of degraded water quality often accompanying the closed state. Cousins et al. \parencite*{cousins_effects_2010} explored seasonal hydrodynamics in Rodeo Lagoon, a eutrophic, infrequently open estuary north of San Francisco, California. The closed condition of Rodeo Lagoon means that tidal forcing is cut off and wind forcing drives circulation and mixing there. Wave overwash provides a salt water source to the lagoon, and resulting density stratification damps turbulence in lower waters and ...\emph{WQ?}... In the summer, wave overwash is rare and wind-mixing converts the lagoon to a well-mixed, brackish state.

Gale et al. \parencite*{gale_vertical_2006} also looked at seasonal timescales in two Australian ICOLLs. \emph{stratification salt dependent but episodic}. 




 Gale et al. ........ in Australia..... \parencite*{gale_vertical_2006}, and Slinger ........ in South Africa \parencite*{slinger_evolution_1990}. Differences between thermal stratification and salinity stratification... 

Measurements or numerical modeling of they hydrodynamics of the transition from an open to closed state (sand bar breach) are essentially non-existant. In the case of a well-mixed lagoon, one would expect the lagoon draining to start as surface flow out and progress quickly to vertically homogenous outflow.  In the case of a highly stratified lagoon, stratification could impede uniform flow out. In the case of a fish mortality event triggered by an artificial mouth opening in the Australian Surrey estuary, it was hypothesized that oxygenated (fresh) surface waters flowed out of the estuary during the breach and hypoxic or anoxic higher salinity water was left behind \parencite{becker_artificial_2009}, though no velocity measurements were made leaving this hypothesis... \emph{(I hate that paper)}. 





%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Where that background leaves us...}

Missing from this work is (a) much of anything in California, (b) short timescales, (c) thinking about estuarine physics in the context of that field which has been defined for larger systems, especially since this field has advanced since measurements were made in the Palmiet estuary in the 1980s... (e.g. dispersion)... in these.. This dissertation aims to begin to fill in some of these holes. 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Observations in the Pescadero estuary in Northern California took place in several field campaigns over 2010 - 2012, and included open state, closed state, and the transitions between these more stable conditions. 

 The Pescadero estuary is just one of many California bar-built estuaries, but one that has been plagued by mortality of federally threatened fish species. One of the goals of this project was to characterize the Pescadero estuary to inform management decisions there. 

Coupled with the current state of understanding, field measurements attempt(ed) to tackle a couple of topics: dynamics of the estuary breach \emph{b/c fish kill}. high frequency forcing \emph{bc obs of IG fluctuations}, and salt dispersion with in the estuary \emph{because it's a thing in bigger systems, trying hard to be a real estuary here}. Furthermore, chance observations during the tsunami generated by the 2011 Tohok$\overline{\mathrm{u}}$ earthquake provided a rare glimse of how small estuaries respond to small tsunamis. 

This thesis will attempt to address four research questions spanning mouth states, timescales, and forcing: 


\subsubsection{Research Questions:}
\begin{enumerate}
	\item{How does the Pescadero estuary function during open, closed, and transitioning mouth states?}
	\item{What are the drivers of hydrodynamics in the open state of small intermittently closed bar-built estuaries?}
	\item{How do tides, stratification, and \emph{marshes} set salt dispersion in a strongly stratified bar-built estuary?}
	\item{How did the tsunami generated by the 2011 Tohok$\overline{\mathrm{u}}$ earthquake alter flow in California's bar built estuaries?}
\end{enumerate}

\emph{to address these questions, the dissertation will follow the outline described below... q1: chapter 2, but also 3-5, q2: ch3, q3: ch 3\&4, q4: ch 5, but built on understanding developed in the previous chapters.}


\subsubsection{Dissertation outline}

In this dissertation, I will present \emph{the latest and greatest understanding...}

Chapter 2 focuses specifically on the Pescadero estuary. In this chapter, the estuary is \emph{described}, field methods relevant to the work in general are presented, and the conditions affecting dynamics in the system in general are described. These are followed by a description of closed state hydrodynamics and a look at how the Pescadero estuary responds to the opening of the closed mouth. This chapter aims to provide insight to land managers through insight into the general functioning of the Pescadero estuary.

In Chapter 3, I examine open-state forcing and hydrodynamics. The connection between waves and nearshore sand transport is well established, but this chapter will look at observed high velocity oscillations caused by this wave forcing as an energetically important variable in California's bar-built estuaries. \emph{sentence on tidally discontinuous flow}. 

Chapters 3 and 4 address salt transport in the open state, first as general movement into and out of the open estuary in Chapter 3, and then in Chapter 4 looking at dispersion of the salt field during the tidal cycle, which varies at different salinities, probably due to interaction with the marsh surrounding the Pescadero estuary.

Finally, Chapter 5 describes the response of the Pescadero estuary to the arrival of waves from the tsunami generated by the 2011 Tohok$\overline{\mathrm{u}}$ earthquake. 

In concluding remarks, Chapter 6 will X. 




%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% END OF CHAPTER ONE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%







reference cite keys:
Cousins: \cite{cousins_effects_2010,cousins_hydrodynamics_2011}
Elwany \cite{elwany_opening_1998}
Behrens \cite{behrens_characterization_2009,behrens_episodic_2013}
Largier \cite{largier_structure_1986,slinger_evolution_1990,largier_circulation_1991,largier_dynamics_1991,largier_stratified_1992,largier_tidal_1992,largier_seasonally_1997}
Rich and Keller \cite{rich_hydrologic_2013}
A. Becker \cite{becker_artificial_2009}
Ranasinghe \cite{ranasinghe_flushing_1998,ranasinghe_circulation_1999,ranasinghe_morphodynamic_1999,ranasinghe_seasonal_2003}
Gale \cite{gale_vertical_2006,gale_processes_2007}
Laudier \cite{laudier_measured_2011} * probably belongs in Pescadero chapter.
Strech (SA) - \cite{stretch_breaching_2006,parkinson_breaching_2007}
ICOLL - \cite{haines_morphometric_2006,baldock_morphodynamic_2008,davidson_simple_2009}
\cite{uncles_infragravity_2014}
Beaches (ICOLL)  \cite{weir_beach_2006}







